%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2012.  
%  For more info about this template or the 
%  TAMU LaTeX User's Group, see http://www.howdy.me/.
%
%  Author: Wendy Lynn Turner 
%	 Version 1.0 
%  Last updated 8/5/2012
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           ACKNOWLEDGEMENTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{ACKNOWLEDGEMENTS}
\addcontentsline{toc}{chapter}{ACKNOWLEDGEMENTS}  % Needs to be set to part, so the TOC doesnt add 'CHAPTER ' prefix in the TOC.

Starting with the obvious, I would like to express my gratitude to Dr. Jim Morel for your guidance and 
willingness to impart your transport wisdom on me over the last few years.  In addition to being a prime source of
knowledge, as an advisor you seem to find the appropriate balance between blind patience and hard-nosed
motivation, where the focus is always exclusively on the problem under investigation.
 I would also like to thank Dr. McClarren and Dr. Ragusa for always willingly providing
alternative view points and discussion on interesting topics, often unrelated to work and
occasionally involving fine beverages. Overall, it has been a pleasure being part of this research group. 

 I would like to thank my parents for a great childhood and continued support, as well
as for never pressuring me to get a real job or take over the farm\ldots yet.  And of course my friends
at A\&M deserve credit for the constant antics and shenanigans that help take the edge off of grad
school.  I must also acknowledge my friend Hank; I cannot imagine my life would have
went down this academic path if we hadn't met.  And finally, I would like to thank my (now) wife Madeline.  I think it goes without much
explanation, but you have always been easy-going through this process and made life pleasant.

Finally, I would like to acknowledge the financial and educational support I have received
over the years. This research was supported with funding received from the DOE Office of Nuclear
Energy's Nuclear Energy University Programs, the DOE National
Nuclear Security Administration, under Award Number(s) DE-NA0002376, and under Los Alamos National Security,
LLC, for the National Nuclear Security Administration of the U.S. Department of
Energy under contract DE-AC52-06NA25396. 
\pagebreak{}
