%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2012.  
%  For more info about this template or the 
%  TAMU LaTeX User's Group, see http://www.howdy.me/.
%
%  Author: Wendy Lynn Turner 
%	 Version 1.0 
%  Last updated 8/5/2012
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           APPENDIX A 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\phantomsection

\chapter{\uppercase{Derivations and Relations for the LO System}}

\section{Useful Moment Relations for LO Equations}
\label{app:lo_mom_relations}

It is useful for derivation and solution of the LO equations to define  several relations between various spatial moments. 
The following relations are derived for $\phi(x)$,
but they can be applied to the moments of general functions.  The cell-averaged $\phi(x)$ 
can be eliminated in terms of the $L$ and $R$ moments using the definition
$b_{L,i}(x)+b_{R,i}(x)=1$ as follows:
\begin{align}\label{app:atoL}
    \phi_{i} &= \frac{1}{h_i} \int\limits_\xl^\xr 1 \;\phi(x) \dd x \\
             &= \frac{1}{h_i} \left(\int\limits_\xl^\xr b_{L,i}(x) \phi(x) \dd x +
             \int\limits_\xl^\xr b_{R,i}(x) \phi(x) \dd x \right)\\
             &= \frac{1}{2} \left( \mom{\phi}_{L,i} + \mom{\phi}_{R,i} \right)
\end{align}
A similar relation between the first moment and the $L$ and $R$ moments as
\begin{equation}
    {\phi}_{x,i} = \frac{3}{2} \left(\mom{\phi}_{R,i} - \mom{\phi}_{L,i}\right)
\end{equation}
The above relations can be inverted to derived a relation for the $L$ and $R$ moments in terms of the slope
and average moments.  These moment expressions are defined purely in terms of integrals,
and are independent of the chosen spatial representation.

Once a linear relation on the interior has been assumed, there are useful spatial closures that can be
derived.  The standard linear interpolatory expansion, for the positive half-range, is restated here:
\begin{equation}
\phi^+(x) = \phi^+_{L,i} b_{L,i}(x) + \phi^+_{R,i} b_{R,i}(x), \quad \xl < x \leq \xr.
\end{equation}
This representation is substituted into the FE basis moment definitions given by
Eq.~\eqref{eq:x_moml} and Eq.~\eqref{eq:x_momr} and the integrals evaluated.  The
resulting expressions produce a relation between the edge values and the outflow from a
cell, i.e., 
\begin{equation}
    \phi_{i,R}^+ = 2 \mom{\phi}_{R,i}^+ - \mom{\phi}_{L,i}^+
\end{equation}
and
\begin{equation}
    \phi_{i,L}^+ = 2 \mom{\phi}_{L,i}^+ - \mom{\phi}_{R,i}^+.
\end{equation}
As in the standard LDFE discretization with upwinding (for $\mu>0$) we make the
approximations $\phi_{i+1/2}^=\phi_{i,R}$ and $\phi_{i-1/2}^+=\phi_{i-1,R}$. Substitution
of the above expressions into the LO equations produce
equations exclusively in terms of moment unknowns.  The solution to these equations produces a linear
representation over each cell that is equivalent to those of a standard LDFE Galerkin
discretization~\cite{fem_book}.

To eliminate the LO unknowns in a manner that produces the same moments as a lumped LDFE
Galerkin discretization, the
following expression can be used for the outflow from a cell
\begin{equation}\label{eq:lump_rep}
    \phi_{i+1/2}^+ = \phi_i^+ + \frac{\phi_{x,i}^+}{3},
\end{equation}
which in terms of the FE moments is equivalent to $\phi_{i+1/2}^+ =
\mom{\phi}_{R,i}^+$.  Inserting this expression into the radiation moment equations, with the same
definition for the linear representation over the interior of $\phi_{i+1/2}^+(x) =
\phi_{L,i} b_{L,i}(x) + \phi_{R,i} b_{R,i}(x)$, will produce an equivalent set of unknowns
as a linear discontinuous method with matrix lumping for the radiation terms.  The
temperature equation and emission source must be independently lumped with an analogous
defintion of the edge values to Eq.~\eqref{eq:lump_rep}. This
relation preserves the average within a cell but does not correctly reproduce the first moment.  

\section{Newtons Method for the LO Equations}
\label{app:lo_newton}

This section briefly derives the equations for the Newton's method solution to the nonlinear LO equations, with the LDFE representation of the temperatures, based on the approach in~\cite{morel_ldtrt}.
Because we have only considered problems with constant densities and heat capacities, the
linearization described below is in terms of temperature $T$ rather than material internal
energy, for simplicity. However, the linearization can be formed in terms of internal energy
to apply this method to a general equation of state.

To formulate the linear equations for each Newton step, the Planckian source is linearized in the material and radiation equations (Eq.~\eqref{eq:mat_td}
\& Eq.~\eqref{eq:trans_td}). 
Application of the first order Taylor expansion in time to the
implicit emission source $\sigma_a a c (T^{n+1})^4$, about some temperature $T^*$ at some
time $t^*\in[t^{n},t^{n+1}]$, yields
\begin{equation}\label{new_planck}
    \sigma_a^{n+1} a c T^{4,n+1} \simeq \sigma_a^* a c \left[T^{*4} + (T^{n+1} - T^*) 4T^{*3} \right]
\end{equation}
where $\sigma_a^*\equiv\sigma_a(T^*)$.  Substitution of this expression into Eq.~\eqref{eq:mat_td} yields
\begin{equation}
    \rho c_v \left( \frac{T^{n+1} - T^{n}}{\Delta t} \right) = \sigma_a^* \phi^{n+1} -
    \sigma_a^* a c \left[ T^{*4} +  (T^{n+1} - T^*) 4T^{*3} \right].
\end{equation}
Algebraic manipulation of this equation yields an expression for $T^{n+1} - T^{*}$:
\begin{align*}
\left( T^{n+1} - T^* \right) &= \frac{ {\displaystyle \frac{\sigma_a^* \Delta t}{\rho
c_v}}  \left[ \phi^{n+1} -  a c T^{*4} \right] + (T^n - T^*) }{1 +
        \sigma_a^* a c \Delta t\frac{\displaystyle 4
T^{*3}}{\displaystyle \rho c_v } }.
\end{align*}
%This provides an expression for $T^{n+1}$ as a
%function of $T^*$ and the mean intensity $\phi^{n+1}$, i.e.,
%\begin{equation}
%\label{lo_t_new}
%T^{n+1}  = \frac{1}{\rho c_v } f\sigma_a^* \Delta t \left( \phi^{n+1} - c a T^{*4} \right)
%+ f T^n + (1-f) T^*.
%\end{equation}
This expression is substituted back into Eq.~\eqref{new_planck} to form
an explicit approximation for the emission source at $t^{n+1}$ as
\begin{equation}\label{t_next1}
    \sigma_a a c T^{4,n+1} \simeq \sigma_a^* (1 -f^*) \phi^{n+1}
    + f^* \sigma_a^* a c T^{4,n} + \rho c_v\frac{1-f^*}{\Delta t} (T^n - T^*)
\end{equation}
where $f^* = [1 + \sigma_a^* c \Delta t 4 a T^{*3}/(\rho c_v)]^{-1}$ is often referred to
as the Fleck factor~\cite{fnc}. 

Next, the above equation is spatially discretized.  Application of the $L$ spatial
moment yields
\begin{multline}\label{eq:temp_const}
    \mom{\sigma_a^* a c T^{4,n+1}}_{L,i} = \sigma^*_{ai}(1-f_i^*)\mom{\phi^{n+1}}_{L,i} +
    f_i^*
    \sigma^*_{ai} a c \left(\frac{2}{3} T_{L,i}^{4,n} + \frac{1}{3} T_{R,i}^{4,n}\right)
    \\ \rho_i c_{vi} \frac{1 - f^*_i}{\Delta t} \left[\frac{2}{3}\left(T^n_{L,i} -
        T^*_{L,i}\right) + \frac{1}{3}\left(T^n_{R,i} -
    T^*_{R,i}\right)\right],
\end{multline}
where $T^{4,n}$ and $T^{n}$ have been assumed LD and $f^*$ is assumed constant over a cell, i.e., $f^*_i
\equiv f(T_i^*)$.
The error introduced by a constant $f^*$ approaches zero as the
non-linearity is converged because $T^*$ approaches $T^{n+1}$. 
Based on an estimate for $T^*$, Eq.~\eqref{eq:temp_const} is an expression for
the Planckian emission source in the radiation moment equations with an additional effective scattering source.
A similar expression can be derived for $\mom{\sigma_{a,i} a c T^4}_R$ and the right
moment equations.
The expressions for the emissions source is substituted into the discrete radiation moment equations,
(Eq.~\eqref{eq:exact_lmomp}--\eqref{eq:exact_rmomm}) to produce a
linear system of equations for the new radiation intensity moments (upon closure of the moment equations). 

Once the linear equations have been solved for new radiation moments, new temperature
unknowns can be estimated.  To conserve energy, the same linearization and discretizations used to
solve the radiation equation must be used in the material energy equation.
Substitution of Eq.~\eqref{eq:temp_const} into the material energy $L$ moment equation, i.e., Eq.~\eqref{eq:lo_mat_dis1},
ultimately yields
\begin{multline}\label{eq:new_temp}
    \frac{2}{3}T_{L,i}^{n+1} + \frac{1}{3}T_{R,i}^{n+1}= \frac{f_i^* \sigma_{ai}^* \Delta
t}{\rho c_{v}}  \left[ \mom{\phi^{n+1}}_{L,i}  - a c\left(\frac{2}{3} T_{L,i}^{4,n} + \frac{1}{3} T_{R,i}^{4,n}\right)
\right] + \\ (1- f^*_{i})\left(\frac{2}{3}T^*_{L,i} + \frac{1}{3}T^*_{R,i}\right) + f
\left(\frac{2}{3}T^n_{L,i} + \frac{1}{3}T^n_{R,i}\right)
\end{multline}
A similar expression is produced for the $R$ moment equation.  This produces a local
matrix equation to solve for new $T_{L,i}$ and $T_{R,i}$ unknowns.  

Based on these equations, iterations on the value of the $T^*$ and FE unknowns for $\phi^{n+1}$ and $T^{n+1}$ can be performed to converge the nonlinearities of the system. The algorithm for solving the LO equations, with iteration index
$m$, is defined as
\begin{enumerate}
    \item Initialize $T^{*,m}$ unknowns using $T^n$ or the last estimate of $T^{n+1}$ from
        previous LO solve.
    \item  Build the LO system based on the effective scattering $(1-f^{*,l})$ and emission terms
        evaluated using $T^{*,m}$.
    \item Solve the linearized LO system to produce a new estimate $\phi^{n+1,m+1}$.
    \item Evaluate a new estimate of $T^{n+1,m+1}$ with energy update equations, e.g., Eq.~\eqref{eq:new_temp}.
       \item $T^{*,m+1}\leftarrow{T}^{n+1,m}$ and $\phi^{n+1,m} \leftarrow
           \phi^{n+1,m+1}$.
    \item Repeat 2-5 until $(T^{n+1,m})^4$ and $\phi^{n+1,m}$ are converged.
\end{enumerate}
Convergence is based on the relative L$_2$ spatial norm of the change in $\phi^{n+1,m}$
and the emission source $\sigma_a a c \left(T^{n+1,m}\right)^4$.

\subsection{Damped Newton Iterations}
\label{app:damped_newton}

The algorithm in the previous section can be modified to improve the stability of
convergence by including a fixed damping factor $\xi$.  In this work, the Newton's method is
formulated to directly estimate the final solution each step, rather that in terms of the
change in the solution between steps.  Thus, an intermediate solve based on the algorithm in the previous section is performed, followed by a damped update of the unknowns. 
The damped Newton's method algorithm is as
follows:
\begin{enumerate}
    \item Choose a damping factor $\xi\in(0,1)$.
    \item Initialize $T^{*,m}$ unknowns using $T^n$ or the last estimate of $T^{n+1}$ from
        previous LO solve.
    \item  Build the LO system based on the effective scattering $(1-f^{*,m})$ and emission terms
        evaluated using $T^{*,m}$.
    \item Solve the linearized LO system to produce an estimate $\phi^{n+1,m+1/2}$.
    \item Evaluate a new estimate of $T^{n+1,m+1/2}$ with energy update equations, e.g., Eq.~\eqref{eq:new_temp}.
    \item Compute new temperatures and intensities as
        \begin{align*}
            \phi^{n+1,m+1} &= \phi^{n+1,m}+ \xi \left( \phi^{n+1,m+1/2} - \phi^{n+1,m} \right)\\
            T^{n+1,m+1} &= T^{n+1,m}+ \xi \left( T^{n+1,m+1/2} - T^{n+1,m} \right).
           \end{align*}
       \item $T^{*,m+1}\leftarrow{T}^{n+1,m}$ and $\phi^{n+1,m} \leftarrow
           \phi^{n+1,m+1}$.
    \item Repeat 2-5 until $(T^{n+1,m})^4$ and $\phi^{n+1,m}$ are converged.
\end{enumerate}

\chapter{\uppercase{Derivations for High-Order Solver}}

\section{Analytic Error Contribution for LDD Trial Space}
\label{sec:face_err_deriv}

In this section, the treatment of the outflow discontinuity residual source and error
tallying is detailed.  Define the additional error contribution from the face
sources at $x_{i+1/2}$ as $\dep$. We have chosen to tally the contribution from these
sources with MC everywhere except for at $x_{i+1/2}$.  Thus, we need to solve for $\dep$
at each face $x_{i+1/2}$ and add that contribution to the tallies
$\epsilon(x_{i+1/2},\mu)$, which include the contribution from all other sources.  The transport equation satisfied by $\dep$, for positive
$\mu$ and effective total cross 
section $\hat \sigma_t$, is
\begin{equation}
    \label{eq:ho_face}
    \mu \pderiv{\dep}{x} + \hat\sigma_t \dep = \rface(x_{i+1/2}^-)\delta^-(x - x_{i+1/2}) + \rface(x_{i+1/2}^+)\delta^+(x - x_{i+1/2}) 
\end{equation}
This equation is integrated from $x_{i+1/2}-\alpha$ to $x_{i+1/2}$ to produce
\begin{multline}
    \mu\dep(x_{i+1/2},\mu) - \mu\dep(x_{i+1/2}-\alpha,\mu)  + \int\limits_{x_{i+1/2}-\alpha}^0 
    \hat \sigma_t \dep \dd x  \\ =  \rface(x_{i+1/2}^-) +
        \int\limits_{x_{i+1/2}-\alpha}^0\rface(x_{i+1/2}^+)\delta^+(x - x_{i+1/2}) \dd x.
\end{multline}
The integral on the right side of the equation is zero because $\delta^+(x-x_{i+1/2})$ is
zero for $(-\infty,x_{i+1/2}]$.  The limit of the above equation is taken as $\alpha\to0$, i.e.,
\begin{multline}
    \lim_{\alpha\to0}\left( \mu\dep(x_{i+1/2},\mu) - \mu\dep(x_{i+1/2}-\alpha,\mu)  + \int\limits_{x_{i+1/2}-\alpha}^0 
    \hat \sigma_t \dep \dd x \right)  = \lim_{\alpha\to0} \rface(x_{i+1/2}^-) 
\end{multline}
The integral goes to zero because $\dep$ is smooth on the interior of the cell, and
$\mu\dep(x_{i+1/2}-\alpha,\mu)$ goes to zero because there is no source upstream of
$x_{i+1/2}^-$. Thus, the final solution is
\begin{equation}
    \dep(x_{i+1/2},\mu) = \frac{\rface(x_{i+1/2}^-)}{\mu} = 
     \tilde I^{(m)}(x_{i+1/2}^-,\mu) - \tilde I^{(m)}(x_{i+1/2},\mu)
.
\end{equation}
The update equation for $I(x_{i+1/2},\mu)$ is thus
\begin{align}
   \tilde I^{(m+1)}(x_{i+1/2},\mu) &= \tilde I^{(m)}(x_{i+1/2},\mu) + \epsilon^{(m)}(x_{i+1/2},\mu) +
    \dep(x_{i+1/2},\mu) \\ 
        &= \tilde I^{(m)}(x_{i+1/2}^-,\mu) + \epsilon^{(m)}(x_{i+1/2},\mu).
\end{align}


\section{Analytic Answer for Fixed Source Problem}
\label{sec:analytic_neutronics}

In this section we model a fixed-source, pure-absorber neutral particle transport
calculation with a known analytic answer and spatial moments of the mean intensity that can be evaluated
semi-analytically.  For a general
isotropic source $q(x)$, the 1D transport equation to be solved is 
\begin{equation}\label{apptr}
    \mu \pderiv{I}{x} + \sigma_a I(x,\mu) = \frac{q(x)}{2}
\end{equation}
A problem is designed to imitate the two-material problem in Sec.~\ref{sec:two} for the
purpose of testing the fix-up for the HO solver in Chapter~\ref{chp:negativities}.  Thus, the domain has two
different constant cross sections and a small isotropic source throughout the domain.  
The boundary conditions are $I(0,\mu)=I_{inc}$ for $\mu>0$ and
$I(X,\mu)=\frac{q(X)}{2\sigma_a}$ for
$\mu<0$, where $x=X$ is the right boundary.  
For positive directions, this first order differential equation is solved using an integration factor
$\exp(\tau(x)/\mu)$ where $\tau(x) = \int_{0}^x \sigma_a(x')\dd x'$.
The solution of Eq.~\eqref{apptr} for the positive half-range intensity is 
\begin{equation}
    I(x,\mu) =  I_{inc}e^{-\tau(x)/\mu} + \int_0^x \frac{q(x')}{2\mu}
    \exp\left(\frac{\ds \tau(x')-\tau(x)}{\ds \mu}\right) \dd x',\quad \mu>0.
\end{equation}
Integration of this result over the positive half range of $\mu$ gives
\begin{equation}\label{e122}
    \phi^+(x) = I_{inc}\E2\left[\tau(x)\right] + \frac{1}{2}\int_0^x q(x')\E1\left[\tau(x)
        - \tau(x')\right]
    \dd x',
\end{equation}
where the linearity of $\tau(x)$ has been used to simplify the argument of the $\E1$
function.

The internal volumetric source is now defined as $q(x) = q_0 \sigma_a(x)$, where $q_0$ is
a constant.  This
definition simulates a floor equilibrium distribution and ensures that $\phi^-(x)$ is a
constant throughout the domain.  This source definition is substituted into
Eq.~\eqref{e122} with the following general relation~\cite{abramowitz}
\begin{equation}
    \E1(y) = -\frac{\dd\E2(y)}{\dd y}  
\end{equation}
to yield a perfect derivative inside the integral. The resulting solution is
\begin{equation}
    \phi^+(x) = I_{inc}\E2\left[\tau(x)\right] + \frac{q_0}{2} \left(1 -
    \E2\left[\tau(x) \right]\right).
\end{equation}
The solution for the negative half range is
\begin{equation}
    \phi^{-}(x) = \frac{q(x)}{2\sigma_a} = \frac{q_0}{2}
\end{equation}
Combination of the above two equations gives the solution for the mean intensity:
\begin{equation}
    \phi(x) = I_{inc}\E2\left[\tau(x)\right] + \frac{q_0}{2} \left(2 -
    \E2\left[\tau(x)\right]\right) .
\end{equation}
This expression can be integrated over any spatial cell, via high-precision quadrature, to accurately approximate moments of
the mean intensity.


\chapter{\uppercase{Derivation of the WLA-DSA Equations}}
\label{sec:wla_derivation}

In this section, we derive the discretized diffusion equation and LD mapping equations
that are used in the WLA-DSA equations.  To simplify notation, we
derive the equations from a generic transport equation (rather than the error equations) with isotropic scattering
and source $q_0$, i.e.,
\begin{equation}\label{eq:ss_trans}
    \mu \pderiv{I}{x} + \sigma_t I = \frac{\sigma_s}{2}\left( \phi(x) + q_0\right).
\end{equation}

\section{Forming a Continuous Diffusion Equation}

First, a continuous spatial discretization of a diffusion equation is derived.  
The mean intensity $\phi$ will ultimately be assumed continuous at faces to produce a
standard three-point finite-difference diffusion discretization. 
The zeroth and first $\mu$ moment of Eq.~\eqref{eq:ss_trans} produce the $P_1$
equations~\cite{lewis,wla_thesis}, i.e., 
\begin{align}\label{eq:dsa_bal}
    \pderiv{J}{x} + \sigma_a \phi &= q_0 \\ \label{eq:p1}
    \sigma_t J + \frac{1}{3} \pderiv{\phi}{x} &= 0.
\end{align}
The spatial finite element moments (defined by Eq.~\eqref{eq:x_moml} and~\eqref{eq:x_momr})
are taken of the above equations. 
The mean intensity is assumed linear on the interior of the cell, i.e.,
$\phi(x)=\phi_Lb_L(x) + \phi_Rb_R(x)$, for $x\in(\xl,\xr)$.   Taking the left moment,
evaluating integrals, and rearranging yields
\begin{equation}
    J_{i} - J_{\il}  + \frac{\sigma_{a,i}h_i}{2} \left(\frac{2}{3} \phi_{L,i} + \frac{1}{3}
    \phi_{R,i} \right) = \frac{h_i}{2} \mom{q}_{L,i}\,\,,
\end{equation}
where $J_i$ is the average of the flux $J$ over the cell. The moments of $q$ are
not simplified to be compatible with the error equations which are in terms of moments. For the $R$ moment
\begin{equation}
    J_{i+1/2} - J_{i}  + \frac{\sigma_{a,i}h_i}{2} \left(\frac{2}{3} \phi_{L,i} + \frac{1}{3}
    \phi_{R,i} \right) = \frac{h_i}{2} \mom{q}_{R,i}\,\,.
\end{equation}
The equation for the $L$ moment is evaluated for cell $i+1$ and added to the $R$ moment
equation evaluated at $i$.  The flux $J$ is assumed continuous at $\ir$ to eliminate
the face fluxes from the equations.  The sum of the two equations becomes
\begin{multline}\label{eq:diff_noclose}
    J_{i+1} - J_{i} + \frac{\sigma_{a,i+1} h_{i+1}}{2}\left(\frac{2}{3} \phi_{L,i+1} +
    \frac{1}{3}\phi_{R,i+1}\right) + \frac{\sigma_{a,i} h_i}{2} \left( \frac{1}{3} \phi_{L,i} +
    \frac{2}{3}\phi_{R,i}\right) =\\ \frac{h}{2} \left(\mom{q}_{L,i+1} + \mom{q}_{R,i}
    \right).
\end{multline}
The mean intensity is approximated as continuous at each face, i.e., $\phi_{L,i+1} = \phi_{R,i}
\equiv \phi_{i+1/2}$.  Adding the $L$ and $R$ moments of Eq.~\eqref{eq:p1} together, with
the continuous approximation for $\phi_{i+1/2}$, produces a discrete Fick's law equation~\cite{stacy}
\begin{equation}\label{eq:ficks}
    J_{i} = -D_i \frac{\phi_{i+1/2} - \phi_{i-1/2}}{h_i},
\end{equation}
where $D_i = 1/(3\sigma_{t,i})$.
Substitution of Eq.~\eqref{eq:ficks} into Eq.~\eqref{eq:diff_noclose} and rearranging yields the following discrete diffusion
equation:
\begin{multline}
        \left(\frac{\sigma_{a,i+1} h_{i+1}}{6} -
        \frac{D_{i+1}}{h_{i+1}}\right)\phi_{i+3/2} + \left(\frac{D_{i+1}}{h_{i+1}} +
        \frac{D_{i}}{h_i} + \frac{\sigma_{a,i+1} h_{i+1}}{3} + \frac{\sigma_{a,i}
        h_{i}}{3}\right)\phi_{i+1/2} \\ + \left(\frac{\sigma_{a,i} h_{i}}{6} -
        \frac{D_{i}}{h_{i}}\right)\phi_{i-1/2} = \frac{h_{i+1}}{2} \mom{q}_{L,i+1} +
        \frac{h_{i}}{2}\mom{q}_{R,i}\;\,. 
\end{multline}
To allow for the use of lumped
or standard LD in these equations, we introduce the factor $\theta$, with
$\theta=1/3$ for standard
LD, and $\theta=1$ for lumped LD.  The diffusion equation becomes
\begin{multline}\label{eq:dsa_lumped_app}
    \left(\frac{\sigma_{a,i+1} h_{i+1}}{4}\left(1 - \theta\right)  -
        \frac{D_{i+1}}{h_{i+1}}\right)\phi_{i+3/2} + \left(\frac{D_{i+1}}{h_{i+1}} +
        \frac{D_{i}}{h_i} + \left(\frac{1+\theta}{2} \right)\left[\frac{\sigma_{a,i+1} h_{i+1}}{2} + \frac{\sigma_{a,i}
        h_{i}}{2}\right]\right)\phi_{i+1/2} \\ + \left(\frac{\sigma_{a,i}
        h_{i}}{4}\left(1 - \theta\right) -
        \frac{D_{i}}{h_{i}}\right)\phi_{i-1/2} = \frac{h_{i+1}}{2} \mom{q}_{L,i+1} +
        \frac{h_{i}}{2}\mom{q}_{R,i}
        \;\,. 
\end{multline}
Summation over all cells forms a system of equations for $\phi$ at each face.  

\subsection{Diffusion Boundary Conditions}

The upwinding in the LO system exactly satisfies the inflow boundary conditions, therefore
a vacuum boundary condition is applied to the diffusion error equations.  The equation for the left moment
at the first cell is given by
\begin{equation}\label{eq:dsa_bc_app}
    J_{1} - J_{1/2}  + \frac{\sigma_{a,i}h_i}{2} \left(\frac{1+\theta}{2} \phi_{L,i}
    + \frac{1-\theta}{2}
    \phi_{R,i} \right) = \frac{h_i}{2} \mom{q}_{L,i}\,\,,
\end{equation}
The Marshak boundary condition for the vacuum inflow at face $x_{1/2}$ is given as
\begin{equation}
    J^+_{1/2} = 0 = \frac{\phi_{1/2}}{4} + \frac{J_{1/2}}{2},
\end{equation}
which can be solved for $J_{1/2}$.  Substitution of the above equation and
Eq.~\eqref{eq:ficks} into Eq.~\eqref{eq:dsa_bc_app} gives 
\begin{equation}\label{eq:bc_dsa_app}
    \left(\frac{1}{2}+ \sigma_{a,1}h_1\frac{1+\theta}{4} - \frac{D_1}{h_1}\right)\phi_{1/2} +
    \left( {\sigma_{a,1}{h_1}}\frac{1-\theta}{4} - \frac{D_1}{h_1}  \right)\phi_{3/2} =
    \frac{h_i}{2} \mom{q}_{L,1}
\end{equation}
A similar expression can be derived for the right-most cell.

\section{Mapping Solution onto LD Unknowns}

Solution of the continuous diffusion equation will provide an approximation to $\phi$ on
faces, denoted as $\phi_{i+1/2}^C$. We now need to map the face solution onto 
the LD representation of $\phi$. To do this, first we take the $L$ and $R$ finite element moments of the P$_1$
equations.  A LDFE dependence is assumed on the interior of the cell for $J$ and
$\phi$.  Taking moments of Eq.~\eqref{eq:dsa_bal} and simplifying yields
\begin{align}
    J_{\ir} - \frac{J_{L,i} + J_{R,i}}{2} + \frac{\sigma_{a,i} h_i}{2} \left(\frac{1}{3} \phi_{L,i} +
    \frac{2}{3}\phi_{R,i}\right) &= \frac{h_i}{2} \mom{q}_{R,i} \\
    \frac{J_{L,i} + J_{R,i}}{2} - J_{i-1/2} + \frac{\sigma_{a,i} h_i}{2}
    \left(\frac{2}{3} \phi_{L,i} +
    \frac{1}{3}\phi_{R,i}\right) &= \frac{h_i}{2} \mom{q}_{L,i}
\end{align}
The moment equations for Eq.~\eqref{eq:p1} are
\begin{align}
    \frac{1}{3}\left(\phi_{\ir} - \frac{\phi_{i,L} + \phi_{i,R}}{2}\right) +
    \frac{\sigma_{t,i} h_i}{2} \left(\frac{1}{3} J_{L,i} + \frac{2}{3}J_{R,i}\right)
    &= 0 \\
    \frac{1}{3}\left(\frac{\phi_{i,L} + \phi_{i,R}}{2} - \phi_{i-1/2} \right) +
    \frac{\sigma_{t,i} h_i}{2} \left(\frac{2}{3} J_{L,i} + \frac{1}{3}J_{R,i}\right)
    &= 0 
\end{align}

The face terms $J_{i\pm 1/2}$ and $\phi_{i\pm 1/2}$ need to be eliminated from the
system. First, the scalar intensity is assumed to be the value provided by the continuous
diffusion solution at each face, i.e., $\phi_{i\pm1/2} = \phi_{i\pm1/2}^C$.
Then, the fluxes are decomposed into half-range values to decouple the equations
between cells.  At $x_{\ir}$, the flux is composed as $J_{i+1/2} = J_{\ir}^+ + J_{\ir}^-$,
noting that in this notation the half-range fluxes are $J_{\ir}^{\pm}=\pm \int_{0}^\pm
\mu I(x_{i+1/2},\mu) \dd \mu$\footnote{Typically, the half-range fluxes are defined with
    integrals weighted with $| \mu |$, but this notation would not be consistent with our
definition of the half-range consistency terms}.  We approximate the incoming fluxes, e.g.,
$J_{i+1/2}^-$, based on $\phi_{i+1/2}^C$ and a P$_1$ approximation.   
The P$_1$ approximation provides the following relation~\cite{wla_thesis}
\begin{equation}
    \phi = 2(J^+ - J^-).
\end{equation}
At $\xr$, the above expression is solved for the incoming current $J_{i+1/2}^-$.  The
total current becomes
\begin{equation}\label{eq:jelim}
    J_{\ir} = J_{\ir}^+ - J_{\ir}^- = 2J_{\ir}^+ - \frac{\phi_{i+1/2}^C}{2},
\end{equation}
In the positive direction, at the right face, the
values of $\phi$ and $J$ are based on the LD representation within the cell at that
face, i.e., $\phi_{R,i}$ and $J_{R,i}$.  The standard P$_1$ approximation for the
half-range fluxes is used\cite{stacy}, i.e.,
\begin{align}
    J^{\pm} &= \frac{\gamma \phi}{2} \pm \frac{J}{2},
\end{align}
where $\gamma$ accounts for the difference between the LO parameters and the true
P$_1$ approximation. Thus, for the right face and positive half-range,
\begin{align}
    J_{\ir}^+ &= \frac{\gamma}{2}\phi_{i,R} + \frac{J_{i,R}}{2} 
\end{align}
A similar expression can be derived for $\xl$.  The total fluxes at each face are
thus
\begin{align}
    J_{i+1/2} &= \gamma\phi_{i,R} + J_{i,R} - \frac{\phi_{\ir}^C}{2} \\
    J_{i-1/2} &= \frac{\phi_{i-1/2}^C}{2} - \gamma \phi_{i,L} + J_{i,L}
\end{align}
Substitution of these results back into the LD balance equations and introduction of the
lumping notation yields the final equations 
\begin{align}\label{eq:update1}
    \left(\gamma\phi_{i,R} + J_{i,R} - \frac{\phi_{\ir}^C}{2} \right) - \frac{J_{L,i} + J_{R,i}}{2} + \frac{\sigma_{a,i} h_i}{2} \left(
    \frac{(1-\theta)}{2} \phi_{L,i} +
    \frac{(1+\theta)}{2}\phi_{R,i}\right) &= \frac{h_i}{2} \mom{q}_{R,i} \\
    \frac{J_{L,i} + J_{R,i}}{2} -\left(\frac{\phi_{i-1/2}^C}{2} - \gamma \phi_{i,L} +
    J_{i,L}\right) + \frac{\sigma_{a,i} h_i}{2} \left(
    \frac{(1+\theta)}{2} \phi_{L,i} +
    \frac{(1-\theta)}{2}\phi_{R,i}\right) &= \frac{h_i}{2} \mom{q}_{L,i} 
    \\
    \frac{1}{3}\left(\phi_{i+1/2}^C - \frac{\phi_{i,L} + \phi_{i,R}}{2}\right) +
    \frac{\sigma_{t,i} h_i}{2}\left( \frac{(1-\theta)}{2} J_{L,i} +
    \frac{(1+\theta)}{2}J_{R,i}\right)    &= 0 \\
    \frac{1}{3}\left(\frac{\phi_{i,L} + \phi_{i,R}}{2} - \phi^C_{i-1/2} \right) +
    \frac{\sigma_{t,i} h_i}{2} \left( \frac{(1+\theta)}{2} J_{L,i} +
    \frac{(1-\theta)}{2}J_{R,i}\right) &= 0 . \label{eq:update2}
\end{align}
The above equations are completely local to each cell and fully defined, including cells
on the boundary. For simplicity, we just take $\gamma=1/2$ for all results.  The system can be solved for the desired unknowns
$\phi_{i,L}$, $\phi_{i,R}$, $J_{i,L}$, and $J_{i,R}$, which represent the mapping of
$\phi_{i+1/2}^C$ onto the LD representation for $\phi^{\pm}(x)$.


